Quantum Mechanics Assignment Three

### 2016-01-27 quantum/latex/a3.tex
finish hw 3

\textbf{Momentum expectation value}
\begin{align*}
  <p> &= -i\sqrt{\frac{\hbar}{2m\omega}}\bra{\psi}\left(a - a^\dagger)\right)\ket{\psi}\\
  &= -i\sqrt{\frac{\hbar m\omega}{2}}\left(\bra{\psi}a\ket{\psi}
  - \bra{\psi}a^\dagger\ket{\psi}\right)\\
  &= -i\sqrt{\frac{\hbar m\omega}{2}}
  \Bigg(
  \sqrt{1}\frac{2}{9}e^{\frac{i\omega t}{2}}e^{\frac{-3i\omega t}{2}}\bra{0}\ket{0}
  + \sqrt{2}\frac{4}{9}e^{\frac{3i\omega t}{2}}e^{\frac{-5i\omega t}{2}}\bra{1}\ket{1}\\
  &-\sqrt{1}\frac{2}{9}e^{\frac{3i\omega t}{2}}e^{\frac{-i\omega t}{2}}\bra{1}\ket{1}
  - \sqrt{2}\frac{4}{9}e^{\frac{5i\omega t}{2}}e^{\frac{-3i\omega t}{2}}\bra{2}\ket{2}
  \Bigg)\\
  &= -i\sqrt{\frac{\hbar m\omega}{2}}
  \left(\frac{2}{9}e^{-i\omega t} + \sqrt{2}\frac{4}{9}e^{-i\omega t}
  - \frac{2}{9}e^{i\omega t} - \sqrt{2}\frac{4}{9}e^{i\omega t}\right)\\
  &= \frac{-2i}{9}\sqrt{\frac{\hbar m\omega}{2}}
  \Bigg(
  \cos(\omega t) - i\sin(\omega t)
  + \sqrt{8}\cos(\omega t) - i\sqrt{8}\sin(\omega t)\\
  &- \cos(\omega t) - i\sin(\omega t)
  - \sqrt{8}\cos(\omega t) - i\sqrt{8}\sin(\omega t)
  \Bigg)\\
  &= \frac{4i}{9}\sqrt{\frac{\hbar m\omega}{2}}\left(1 + \sqrt{8}\right)\sin(\omega t)\\
\end{align*}

@@

@@

We begin with the perturbed matrix and diagonalize it:\\

\begin{align*}
  H &\doteq
  \pm\frac{\hbar}{2}
  \begin{pmatrix}
    \omega_0 + \omega_1 & \omega_2\\
    \omega_2 & -\omega_0 - \omega_1\\
  \end{pmatrix}
  =
  \pm\frac{\hbar}{2}
  \begin{pmatrix}
    -\omega_2 & \omega_2\left(\frac{-\omega_2}{\omega_0 + \omega_1}\right)\\
    \omega_2 & -\omega_0 - \omega_1\\
  \end{pmatrix}
  =
  \pm\frac{\hbar}{2}
  \begin{pmatrix}
    -\omega_2 & \frac{-\omega_2^2}{\omega_0 + \omega_1}\\
    0 & -\omega_0 - \omega_1 - \frac{-\omega_2^2}{\omega_0 + \omega_1}\\
  \end{pmatrix}\\
  &=
  \pm\frac{\hbar}{2}
  \begin{pmatrix}
    -\omega_2 & \frac{-\omega_2^2}{\omega_0 + \omega_1}\\
    0 & \frac{\omega_2^2}{\omega_0 + \omega_1}\\
  \end{pmatrix}
  =
  \pm\frac{\hbar}{2}
  \begin{pmatrix}
    -\omega_2 & 0\\
    0 & \frac{\omega_2^2}{\omega_0 + \omega_1}\\
  \end{pmatrix}
\end{align*}

>>>>

\textbf{Momentum expectation value}
\begin{align*}
  <p> &= -i\sqrt{\frac{\hbar}{2m\omega}}\bra{\psi}\left(a - a^\dagger)\right)\ket{\psi}\\
  &= -i\sqrt{\frac{\hbar m\omega}{2}}\left(\bra{\psi}a\ket{\psi}
  - \bra{\psi}a^\dagger\ket{\psi}\right)\\
  &= -i\sqrt{\frac{\hbar m\omega}{2}}
  \Bigg(
  \sqrt{1}\frac{2}{9}e^{\frac{i\omega t}{2}}e^{\frac{-3i\omega t}{2}}\bra{0}\ket{0}
  + \sqrt{2}\frac{4}{9}e^{\frac{3i\omega t}{2}}e^{\frac{-5i\omega t}{2}}\bra{1}\ket{1}\\
  &-\sqrt{1}\frac{2}{9}e^{\frac{3i\omega t}{2}}e^{\frac{-i\omega t}{2}}\bra{1}\ket{1}
  - \sqrt{2}\frac{4}{9}e^{\frac{5i\omega t}{2}}e^{\frac{-3i\omega t}{2}}\bra{2}\ket{2}
  \Bigg)\\
  &= -i\sqrt{\frac{\hbar m\omega}{2}}
  \left(\frac{2}{9}e^{-i\omega t} + \sqrt{2}\frac{4}{9}e^{-i\omega t}
  - \frac{2}{9}e^{i\omega t} - \sqrt{2}\frac{4}{9}e^{i\omega t}\right)\\
  &= \frac{-2i}{9}\sqrt{\frac{\hbar m\omega}{2}}
  \Bigg(
  \cos(\omega t) - i\sin(\omega t)
  + \sqrt{8}\cos(\omega t) - i\sqrt{8}\sin(\omega t)\\
  &- \cos(\omega t) - i\sin(\omega t)
  - \sqrt{8}\cos(\omega t) - i\sqrt{8}\sin(\omega t)
  \Bigg)\\
  &= \frac{-4}{9}\sqrt{\frac{\hbar m\omega}{2}}\left(1 + \sqrt{8}\right)\sin(\omega t)\\
\end{align*}

@@

This was a cool problem that shows that it is very easy to construct a Hamiltonian for a given
quantum system in matrix representation (not necessarily position representation). It is
important to note that operator A is not an actual physical operator, since it has no eigenvalue
equation.\\

\clearpage

@@

We begin with the perturbed matrix and use the fact that the eigenvalues of a matrix will be the
diagonal elements in the diagonalized form:\\

\begin{align*}
  H &\doteq
  \pm\frac{\hbar}{2}
  \begin{pmatrix}
    \omega_0 + \omega_1 & \omega_2\\
    \omega_2 & -\omega_0 - \omega_1\\
  \end{pmatrix}\\
\end{align*}

\begin{align*}
  \det\left(H - \lambda I\right) &= 0\\
  \left(\omega_0 + \omega_1 - \lambda\right)
  \left(-\omega_0 - \omega_1 - \lambda\right) - \omega_2^2 &= 0\\
  \lambda^2 - \left(\omega_0 + \omega_1\right)^2 - \omega_2^2 &= 0\\
  \lambda = \pm\sqrt{\left(\omega_0 + \omega_1\right)^2 + \omega_2^2}\\
\end{align*}

We then use these eigenvalues to construct a diagonal matrix. We know that the basis for the
perturbed Hamiltonian will be very similar to the unperturbed Hamiltonian, so we arrange our
perturbed Hamiltonian with the top left element positive and bottom right negative:

\begin{align*}
  H_p &\doteq
  \pm\frac{\hbar}{2}
  \begin{pmatrix}
    \sqrt{\left(\omega_0 + \omega_1\right)^2 + \omega_2^2} & 0\\
    0 & -\sqrt{\left(\omega_0 + \omega_1\right)^2 + \omega_2^2}\\
  \end{pmatrix}\\
\end{align*}

We have already shown that the eigenvalues agree with those in McIntyre (10.9). What we have done in
this problem is taken a matrix in the unperturbed energy basis and transformed it into its own
perturbed basis. We know this is a new basis because the perturbed Hamiltonian in the original
basis was not diagonal.\\

It is a good confirmation that the method used in McIntyre to find the eigenvalues of the perturbed
system, redefining the direction of Z to point in the direction of the new B-field, yielded the same
result as creating a new basis for the perturbed Hamiltonian.\\
